---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ACHSproj)
```

# ACHSproj

Companion package for the ACHS Spring 2026 talk. It provides tools for
projecting payout annuity cashflows, computing present values and IRRs, and
visualising results — all built on S7 classes and the vctrs/tidyverse
ecosystem.

## Projecting a payout annuity

`annuity_cf()` is the core function. It builds a monthly projection for a
deferred, life-contingent payout annuity and returns a `projection` object
whose `@data` tibble contains mortality, cashflow, and reserve columns.

Key assumptions in this projection:

- Mortality follows the 2012 Individual Annuity Mortality Basic table
- Mortality improvement follows Projection Scale G2 with improvements from 2012
- A single deterministic discount rate
- Monthly cashflows

```{r annuity-example}
p <- annuity_cf(
  issue_age    = 60,
  benefit      = 8,       # annual benefit amount
  premium      = 100,     # single premium at time 0
  defer_years  = 2L,      # 2-year deferral period
  certain_years = 3L,     # 3-year certain period
  refund       = TRUE,    # return-of-premium death benefit
  disc         = 0.05,    # annual discount rate
  start_date   = "2026-05-06"
)

p
```

Key columns in `p@data`:

| Column | Description |
|---|---|
| `tpx` | Cumulative survival probability from time 0 |
| `premium` | Single premium inflow at time 0 |
| `certain_cf` | Guaranteed benefit payments during the certain period |
| `life_cf` | Life-contingent benefit payments after the certain period |
| `death_cf` | Return-of-premium death benefit (when `refund = TRUE`) |
| `total_cf` | Sum of all cashflows |
| `reserve` | Prospective reserve — PV of future cashflows at each period |

**NOTE:** The `reserve` column is for illustrative purposes only and is not meant to represent a realistic reserve that would be held under any particular regulatory regime.

## Support functions

### `pv()` — present value

`pv()` works on both plain numeric vectors and `projection` objects.

```{r pv-vector}
# PV of a simple cash flow stream: invest 100 today, receive 110 in one year
pv(c(-100, 110), disc = 0.05, freq = 1L)
```

On a `projection`, it returns a tibble of present values for every `cf_vec`
column, discounted at the stored rate by default.

```{r pv-projection}
pv(p)
```

The `rolling = TRUE` option returns the prospective reserve at every time step
— that is, the PV of cashflows still to come, discounted back to that moment.
This is what populates the `reserve` column automatically.

```{r pv-rolling}
head(pv(p, rolling = TRUE)$total_cf, 10)
```

### `irr()` — internal rate of return

`irr()` finds the discount rate at which the PV of a cashflow stream is zero.

```{r irr-vector}
# Invest 100, receive 110 in one year — IRR should be 10%
irr(c(-100, 110), freq = 1L)
```

On a `projection`, `irr()` operates on `total_cf` and returns the **cost of
funds** — the rate at which the insurer breaks even on the contract.

```{r irr-projection}
irr(p)
```

### `plot()` — cashflow and reserve chart

`plot()` produces a two-panel chart: the top panel shows the cashflow
components over time, and the bottom panel shows the prospective reserve.

```{r plot, fig.alt = "Two-panel line chart of payout annuity cashflows and reserve"}
plot(p)
```
